---
title: "Initial Korean covid 19"
author: "Jamal Evans"
date: "spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Introduction

$H_0$ is that there is no difference in death rate based on sex.

$H_A$ is that there IS a difference in death rate

# Methods
variables for this study are comprised of Sex and result 'dead' or 'alive'

We want to know if result depends on Sex
Result is a factor variable (Categorical), while Sex is also a factor Categorical variable.
This means we use cat~cat helper file.

## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This Bar Graph shows that males are more likely to die according to percentage 60 percent male 40 percent female.
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The Numerical Data shows higher percentage of death in males from 64 percent to 36 for females.
## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-squared test (p-value= 0.0012) shows that the new p-value is more than 0.05, meaning we reject the null
This means there IS a significant difference in death rate based on Sex.

The Fisher exact test (p-value=0.0009) s.
The odds ratio is 2.3 and is not equal to 1 this means that we reject the null ($H_0$).

## Conclusions
The original question was "is there a difference in death rate in the Korean data set based on Sex".
The data plot shows that there is a considerable difference in death higher for males than females.
The numerical result confirmed the percentage from 1.3 to 3 in death rating. In the end we reject the null
